Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-08-17 writeup/rocha_etal_supporting-info.tex
Adds reference to ECCO project
Syncronize renaming
Fixes typo in SI
Updates supporting information

\noindent\textbf{Contents of this file}
%%%Remove or add items as needed%%%
\begin{enumerate}
\item Text S1 with details of the LLC simulations.
\item Figure S1 with spectra of LLC2160 and LLC4320 simulations.
\item Table S1 with details about the LLC spin-up.
\item Figure S2 with time series of vorticity and divergence probability density
      functions.
\item Figure S3: root-mean-squares of vorticity, strain rate, and divergence
      as a function of depth.
\item Figure S4 with comparison of Argo climatological upper-ocean stratification
      and the LLC simulations
\item Text S5 with details about the comparison of model and mooring kinetic energy
            frequency spectra.
\item Figure S5: comparison of upper-ocean kinetic energy
      frequency spectra with spectra from available moored current meters.
\item Figure S6 showing the seasonal variations of the near-surface amplitude
      of the vertical modes.
%if Tables are larger than 1 page, upload as separate excel file
\end{enumerate}

@@

Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC code and model configuration files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

@@

 \begin{figure}
    \begin{center}
      \setfigurenum{S1}
      \includegraphics[width=40pc]{figs/fig_s2_3.pdf}
   \caption{Comparison between spectra of the 1/24$^\circ$
   (green) and 1/48$^\circ$ (purple) LLC simulations. The statistical error bars represent conservative
   95$\%$ significance levels.  As expected the
   higher-resolution simulation (LLC4320) essentially extends the spectra of the
   lower-resolution simulation (LLC2160) towards smaller scales, demonstrating that
   the higher resolution is resolving a wider array of scales. Between 20 and 100 km
   the spectra of the two simulations are visually indistinguishable.}
   \label{figS2_3}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S2}
      \includegraphics[width=40pc]{figs/fig_s2_2.pdf}
   \caption{Time dependence of probability density functions of vertical vorticity (a, b)
   and horizontal divergence (c, d). Monthly averages in April and October are
   also shown (e, f). The time series of vorticity PDF indicate that
   the development of positive skewness in winter (Figures \ref{figS2_2}a-b) both in
   hourly ($\sim 1.4$) and daily-averaged ($\sim1.1$) fields. The skewness
   in divergence for hourly fields is very small year-round ($\sim -0.24$), suggesting the
   prevalence of inertia-gravity waves. Filtering out the waves by daily-averaging
   the velocity field increases the skewness in divergence, particularly in winter ($\sim-0.64$). Both
  positive vorticity (cyclonic) skewness and negative divergence (convergence) skewness
  are characteristics of submesoscale turbulence.}
   \label{figS2_2}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S3}
      \includegraphics[width=40pc]{figs/fig_s2_2.png}
   \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
   and horizontal divergence (c,f) for hourly (a,b,c) and daily-averaged (d,e,f,) fields.
   The strong seasonality, as depicted
   at the surface in Figure 2 of the main manuscript, is confined to the mixed layer, which
   varies from 50 m in summer to about 350 m in winter.}
   \label{figS2_2}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S4}
      \includegraphics[width=40pc]{figs/fig_s2_1.pdf}
   \caption{Potential density: a comparison between Roemmich-Gilson Argo
           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
           and  1/48$^\circ$ (e, f) LLC simulations. Of course, the Argo climatology is smoother,
           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce the strong
           seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
           and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
            during all seasons (not shown). The Argo climatology is available online at
           \texttt{http://sio-argo.ucsd.edu/RG\_Climatology.html}.}
   \label{figS2_1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S5}
      \includegraphics[width=40pc]{figs/figA1.pdf}
   \caption{Comparison between frequency spectra of LLC simulations and two
   available moored current meter records.}
   \label{figA1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S6}
      \includegraphics[width=40pc]{figs/fig_s3.pdf}
   \caption{The seasonal variability of the WOA 2013 \citep{levitus_etal2013} stratification averaged  over
   the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.
   Clearly, the changes to the stratification
   are dramatic, with the formation of a seasonal pycnocline in summer. The strong
   seasonality of upper-ocean stratification yields a strong seasonality in the
   near-surface shape and amplitude of baroclinic pressure modes.
   In particular, the surface amplitude is much larger in summer. Thus the projection
   of baroclinic tides on the surface may have significant seasonal variations. The WOA data is
   available online at \texttt{https://www.nodc.noaa.gov/OC5/woa13/}.}
   \label{figS3}
   \end{center}
 \end{figure}

>>>>

\noindent\textbf{Contents of this file}
%%%Remove or add items as needed%%%
\begin{enumerate}
\item Text S1 with details of the LLC simulations.
\item Figure S1 with spectra of LLC2160 and LLC4320 simulations.
\item Table S1 with details about the LLC spin-up.
\item Figure S2 with time series of vorticity and divergence probability density
      functions.
\item Figure S3: root-mean-squares of vorticity, strain rate, and divergence
      as a function of depth.
\item Figure S4 with comparison of Argo climatological upper-ocean stratification
      and the LLC simulations
\item Text S5 with details about the comparison of model and mooring kinetic energy
            frequency spectra.
\item Figure S5: comparison of upper-ocean kinetic energy
      frequency spectra with spectra from available moored current meters.
\item Figure S6 showing the seasonal variations of the near-surface amplitude
      of the vertical modes.
\item Figure S7 Wavenumber-frequency spectrum of KE for LLC4320 simulation.
%if Tables are larger than 1 page, upload as separate excel file
\end{enumerate}

@@

Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC code and model configuration files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). The model output can be obtained from  obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}).

@@

 \begin{figure}
    \begin{center}
      \setfigurenum{S1}
      \includegraphics[width=40pc]{figs/figS1.pdf}
   \caption{Comparison between spectra of the 1/24$^\circ$
   (green) and 1/48$^\circ$ (purple) LLC simulations. The statistical error bars represent conservative
   95$\%$ significance levels.  As expected the
   higher-resolution simulation (LLC4320) essentially extends the spectra of the
   lower-resolution simulation (LLC2160) towards smaller scales, demonstrating that
   the higher resolution is resolving a wider array of scales. Between 20 and 100 km
   the spectra of the two simulations are visually indistinguishable.}
   \label{figS2_3}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S2}
      \includegraphics[width=40pc]{figs/figS2.pdf}
   \caption{Time dependence of probability density functions of vertical vorticity (a, b)
   and horizontal divergence (c, d). Monthly averages in April and October are
   also shown (e, f). The time series of vorticity PDF indicate that
   the development of positive skewness in winter (Figures \ref{figS2_2}a-b) both in
   hourly ($\sim 1.4$) and daily-averaged ($\sim1.1$) fields. The skewness
   in divergence for hourly fields is very small year-round ($\sim -0.24$), suggesting the
   prevalence of inertia-gravity waves. Filtering out the waves by daily-averaging
   the velocity field increases the skewness in divergence, particularly in winter ($\sim-0.64$). Both
  positive vorticity (cyclonic) skewness and negative divergence (convergence) skewness
  are characteristics of submesoscale turbulence.}
   \label{figS2_2}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S3}
      \includegraphics[width=40pc]{figs/figS3.png}
   \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
   and horizontal divergence (c,f) for hourly (a,b,c) and daily-averaged (d,e,f,) fields.
   The strong seasonality, as depicted
   at the surface in Figure 2 of the main manuscript, is confined to the mixed layer, which
   varies from 50 m in summer to about 350 m in winter.}
   \label{figS2_2}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S4}
      \includegraphics[width=40pc]{figs/figS4.pdf}
   \caption{Potential density: a comparison between Roemmich-Gilson Argo
           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
           and  1/48$^\circ$ (e, f) LLC simulations. Of course, the Argo climatology is smoother,
           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce the strong
           seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
           and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
            during all seasons (not shown). The Argo climatology is available online at
           \texttt{http://sio-argo.ucsd.edu/RG\_Climatology.html}.}
   \label{figS2_1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S5}
      \includegraphics[width=40pc]{figs/figS5.pdf}
   \caption{Comparison between frequency spectra of LLC simulations and two
   available moored current meter records.}
   \label{figA1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S6}
      \includegraphics[width=40pc]{figs/figS6.pdf}
   \caption{The seasonal variability of the WOA 2013 \citep{levitus_etal2013} stratification averaged  over
   the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.
   Clearly, the changes to the stratification
   are dramatic, with the formation of a seasonal pycnocline in summer. The strong
   seasonality of upper-ocean stratification yields a strong seasonality in the
   near-surface shape and amplitude of baroclinic pressure modes.
   In particular, the surface amplitude is much larger in summer. Thus the projection
   of baroclinic tides on the surface may have significant seasonal variations. The WOA data is
   available online at \texttt{https://www.nodc.noaa.gov/OC5/woa13/}.}
   \label{figS3}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S7}
      \includegraphics[width=40pc]{figs/figS7.pdf}
   \caption{LLC4320 Wavenumber-frequency spectrum of KE in (a) April
    and (b) October. (c) Wavenumber spectrum of SSH variance — the integral of (a)
    and (b) over frequency. In (a) and (b), the light gray shaded region depicts the
    dispersion relations for inertia-gravity waves from mode 1 through mode 4 across
    the latitudinal domain; horizontal dashed lines represent the semi-diurnal lunar
    tidal frequency (M2), and the inertial frequency at mid-latitude (f32.5) and its
    first harmonic (2f32.5).}
   \label{figS7}
   \end{center}
 \end{figure}
